Luminance constancy under fixed geometry

### 2018-03-24 Draft2/draft2.tex
Luminous reflectance factor

We define the computational problem of luminance constancy as that of estimating what we call surface luminance. 
Surface luminance is a measure of the overall amount of light that a surface reflects.
We define

A surface' luminance is the luminance of the light that would be reflected to the eye from a flat matte object with that surface reflectance, when the object is diffusely illuminated by the standard CIE daylight spectrum D65 \cite{CIE86}.  We then compute the luminance of this reflected light, using the CIE 1931 photopic luminosity function \cite{CIE86}, and divide this by the luminance of a perfectly reflecting diffuser (surface reflectance of one at all wavelengths) when that is computed in the same way. This procedure allows us to assign to any surface reflectance a single surface luminance value.  Because the perceived luminance of an object is approximately (but by no means exactly, see \citeNP{wyszecki1986color}) correlated with surface luminance across a range of viewing conditions, we refer to the computational problem we study as that of {\em luminance constancy}.
This is a generalization of the way this term is often used, where it refers more specifically to the stability of perceived luminance across images of scenes where all objects and illuminants are achromatic \cite{gilchrist2006seeing}.
Here, we aim to estimate surface luminance, but do so for images where both the luminance and the chromaticity of the surfaces and illuminants varies (see  Fig.~\ref{fig:introExampleFigure}). A particularly important feature of our treatment is that we consider scene variations where not only the spectrum of the illumination varies, but also the surface reflectance of objects in the scene other than the object whose surface luminance is being estimated.

% MOVE ALL THIS TO DICUSSION
%This general approach has been adopted by other labs for the study of computational estimation of optical flow (Baker et al., 2011) and is becoming increasingly popular (Kovacs, Bell, Snavely, & Bala, 2017), as well as for the generation of stimuli for the study of human perception (e.g., Boyaci, Maloney, & Hersh, 2003; Todd & Mingolla, 1983; Arend & Reeves, 1986; Johnston, Cumming, & Parker, 1993). Of course, success within the domain of graphics images does not guarantee smooth generalization to real natural images, a point we return to in the Discussion.
% \cite{baker2011database}  \cite{kovacs17shading} \citeNP{boyaci2003effect, todd1983perception, arend1986Simultaneous, johnston1993integration}
% Autonmous vehicle work - ask BAW

 {\em virtual natural scenes}.

 we also model the initial steps of the human visual system, so that our supervised learning methods are applied not to image pixels but rather to modeled responses of the foveal cone photoreceptor mosaic.

Even within the special case of luminance constancy, there are many scene manipulations that could be considered. The set of manipulations we focus on here are illustrated by Fig.~\ref{fig:introExampleFigure} and the three panels of Fig.~\ref{fig:studiedCases}. These cases are: 1. variation in target object spectrum with fixed light source and background objects spectra (Fig.~\ref{fig:introExampleFigure}), 2. variation in target object spectrum and background objects spectra, with fixed light source  (Fig.~\ref{fig:backgroundVarying}), 3. variations in both the target object spectrum and the light source spectra, with fixed background object spectra (Fig.~\ref{fig:targetIlluminantVarying}), and 3. variations in all three types of spectra (target object, light source, and background objects) (Fig.~\ref{fig:allSpectraVarying}). Across real scenes, there will also be variations in the shape of the target object, its pose, eye position, etc. (see \nameref{Methods}). In this paper, we focus on the problem of luminance constancy within scene ensembles where these geometric factors are held fixed.

We show that for cases that have spectral variation only either in target object surface reflectance or background object surface reflectance or the power spectrum of the light sources, luminance can be estimated by simple transformations of the cone responses. For variations only in target object reflectance spectra (case 1) or just the target and background objects reflectance spectra (case 2), object luminance can be recovered directly from the light reflected off the target object. For case 3, where both the target object reflectance and light source spectra vary, object luminance can be recovered from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 4), the standard luminance can not be recovered simply from the cone response or the contrast. We have used AMA with a Bayesian estimator to estimate standard luminance of the target object. Using this method, we can recover the luminance of the target object with $\sim 15\%$ relative root mean square error.  Our analysis shows that the optimal receptive fields for estimating the luminance of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, luminance estimation involves a weighted sum of the L, M and S cone responses where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performance of the RFs learnt using  one case on the cone response of other cases shows that the receptive field of the complex cases perform well on images of simpler cases. We discuss the implications of such receptive fields on luminance constancy.

The rest of this paper is organized as follows. In the \nameref{Methods} section, we describe in detail the process to generate naturalistic labeled images (\nameref{method:VirtualWorld}) using our software pipeline {\rm Virtual world color constancy}, the process to simulate the response of retinal cone photoreceptors to these images using a standard model of the early visual system (\nameref{method:Isetbio}) and the supervised learning method (AMA) used in our analysis (\nameref{method:SupervisedLearning}). In the \nameref{Results} section, we show the estimates and performance of the supervised methods for the four cases mentioned above. We conclude the paper with the implication of our results in the \nameref{Discussion} section.

>>>>

We define the computational problem of luminance constancy as that of estimating the luminous reflectance factor of the diffuse component of an object's surface reflectance function.
The luminous reflectance factor is a measure of the overall amount of light reflected by a matte surface.
More precisely, the luminous reflectance factor is the normalized luminance of the light that would
be reflected to the eye from a flat matte object with that the specified surface reflectance function, when the object is diffusely illuminated by a reference illuminant.
The normalization factor is 100 divided by the luminance of the reference illuminant, so that luminous reflectance factors range between 0 and 100.
Here we use CIE daylight D65 as the reference illuminant and the CIE 1931 photopic luminosity function to compute luminance  \cite{CIE86}.

% MOVE THIS TO DICUSSION
%This general approach has been adopted by other labs for the study of computational estimation of optical flow (Baker et al., 2011) and is becoming increasingly popular (Kovacs, Bell, Snavely, & Bala, 2017), as well as for the generation of stimuli for the study of human perception (e.g., Boyaci, Maloney, & Hersh, 2003; Todd & Mingolla, 1983; Arend & Reeves, 1986; Johnston, Cumming, & Parker, 1993). Of course, success within the domain of graphics images does not guarantee smooth generalization to real natural images, a point we return to in the Discussion.
% \cite{baker2011database}  \cite{kovacs17shading} \citeNP{boyaci2003effect, todd1983perception, arend1986Simultaneous, johnston1993integration}
% Autonmous vehicle work - ask BAW

% NOT SURE WE NEED ANY OF THIS.  PROBABLY NOT HERE IF WE DO
%Even within the special case of luminance constancy, there are many scene manipulations that could be considered. The set of manipulations we focus on here are illustrated by Fig.~\ref{fig:introExampleFigure} and the three panels of Fig.~\ref{fig:studiedCases}. These cases are: 1. variation in target object spectrum with fixed light source and background objects spectra (Fig.~\ref{fig:introExampleFigure}), 2. variation in target object spectrum and background objects spectra, with fixed light source  (Fig.~\ref{fig:backgroundVarying}), 3. variations in both the target object spectrum and the light source spectra, with fixed background object spectra (Fig.~\ref{fig:targetIlluminantVarying}), and 3. variations in all three types of spectra (target object, light source, and background objects) (Fig.~\ref{fig:allSpectraVarying}). Across real scenes, there will also be variations in the shape of the target object, its pose, eye position, etc. (see \nameref{Methods}). In this paper, we focus on the problem of luminance constancy within scene ensembles where these geometric factors are held fixed.
%
%We show that for cases that have spectral variation only either in target object surface reflectance or background object surface reflectance or the power spectrum of the light sources, luminance can be estimated by simple transformations of the cone responses. For variations only in target object reflectance spectra (case 1) or just the target and background objects reflectance spectra (case 2), object luminance can be recovered directly from the light reflected off the target object. For case 3, where both the target object reflectance and light source spectra vary, object luminance can be recovered from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 4), the standard luminance can not be recovered simply from the cone response or the contrast. We have used AMA with a Bayesian estimator to estimate standard luminance of the target object. Using this method, we can recover the luminance of the target object with $\sim 15\%$ relative root mean square error.  Our analysis shows that the optimal receptive fields for estimating the luminance of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, luminance estimation involves a weighted sum of the L, M and S cone responses where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performance of the RFs learnt using  one case on the cone response of other cases shows that the receptive field of the complex cases perform well on images of simpler cases. We discuss the implications of such receptive fields on luminance constancy.
